Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
Udpates

\begin{figure}[t!]
	\centering
	\includegraphics[page=1]{fig5.pdf}
	\caption{DAG learned from the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert. The p-value threshold used is $ 0.05 $ and the measure of association threshold is $ 0. 1 $}
	\label{fig:adult_llm}
\end{figure}

@@

To enable users to apply this method to their own datasets, we developed an
interactive web-based tool (shown in Fig.~\ref{fig:web}) for constructing
models. Users begin by uploading their dataset, which initializes an empty DAG
with nodes corresponding to the dataset\'s variables. They can then specify a
p-value threshold and a measure of association threshold. To make the method
work for mixed data types, we use a measure based on canonical correlations
(Appendix~\ref{sec:mixed_association}).

Using the specified thresholds, the tool visually highlights unexplained
correlations by displaying red edges between variables where correlation exists
in the data but the current DAG is not able to explain. The thickness of these
edges represents the strength of the correlation, helping users prioritize
which edges to add. Similarly, if an edge in the graph is detected to be
unnecessary, it is highlighted in black. Based on this information, users can
select to remove unnecessary edge or select a potential edge to add and specify
the orientation of the edge. The tool computes Shipley’s C \citep{Shipley2000}
value at each change to determine the overall fit of the model to the data.
Once satisfied with the constructed DAG, users can export the model for further
analysis.

>>>>

\begin{figure}[t!]
	\centering
	\includegraphics[page=1]{fig5.pdf}
	\caption{DAG learned from the Adult Income dataset using an LLM
		(Gemini 1.5 Flash) as the expert. The p-value threshold used is $ 0.05 $ 
		and the measure of association threshold is $ 0. 1 $}
	\label{fig:adult_llm}
\end{figure}

@@

To enable researchers to easily apply our approach to their own datasets, we
also developed an interactive web-based tool (shown in Fig.~\ref{fig:web}) for
constructing models. Users can upload their dataset, which initializes an empty
DAG with nodes corresponding to the dataset's variables. They can then specify
a p-value threshold and a measure of association threshold. To make the method
work for mixed data types, we use a measure based on canonical correlations
(Appendix~\ref{sec:mixed_association}).

Using the specified thresholds, the tool visually highlights variable pairs
(using red edges) with a residual association higher than the threshold. The
thickness of these edges represents the strength of association, helping users
prioritize which variable pairs to address first. Similarly, if an edge in the
graph is detected to be unnecessary, it is highlighted in black. Based on this
information, make modifications to the model by adding or removing edges. The
tool also provides the Room Mean Square Error of Approximation (RMSEA) of the
overall model fit to the data. Once satisfied with the constructed DAG, users
can export the model for further analysis.
